---
title: "PREMS Analyse de données ANT"
author: "Christophe Pallier"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, ouput="hide", message=FALSE, warning=FALSE}
require(readr)
require(stringr)
require(plyr)
require(dplyr)
require(ggplot2)
require(tidyr)
require(ez)
```


The control group data are in `0T_behav_data` and the exeprimental group is in `11.7T_behav_data` (There are 4 `*.xpd` files per subject corresponding to each of the 4 runs of the ANT task: 2 ran outside of the scanner on pc6, and 2 ran inside the scanner on pc11).


# Import Data 

```{r, echo=FALSE, ouput="hide", message=FALSE, warning=FALSE}
setwd("ant_data")
xpd_files <- list.files(path='.', pattern = "*.xpd")
alld <- readr::read_csv(xpd_files, id="file_name", comment="#", na="None")
setwd("..")

n_subj = length(unique(alld$subject_id))
files <- sort(unique(alld$file_name))
stopifnot(length(files) == n_subj * 4) 

for (i in 1:n_subj)
{
   index = (i - 1) * 4
   alld <- alld %>% mutate(file_name = str_replace(file_name, files[1 + index], "1_out")) %>% 
                    mutate(file_name = str_replace(file_name, files[2 + index], "2_in")) %>%
                    mutate(file_name = str_replace(file_name, files[3 + index], "3_in_noise")) %>%
                    mutate(file_name = str_replace(file_name, files[4 + index], "4_out")) 
}

alld <- alld %>% rename("run" = "file_name")
alld$run = as.factor(alld$run)
alld$subject_id = as.factor(alld$subject_id)
alld$cue_validity = as.factor(alld$cue_validity)
alld$flanker_congruency <- as.factor(alld$flanker_congruency)

alld <- mutate(alld, "group" = factor(ifelse(as.integer(as.character(alld$subject_id)) > 21, "experimental", "control")))  # from subj22, we are in the experimental group
alld <- mutate(alld, "correct" = ifelse(alld$run %in% c("1_out", "4_out"), 
       (alld$response_key==102 & alld$arrow_direction=='left') | (alld$response_key==106 & alld$arrow_direction=='right'),
       (alld$response_key==121 & alld$arrow_direction=='left') | (alld$response_key==98 & alld$arrow_direction=='right')))
alld$correct[is.na(alld$response_key)] = FALSE
```

Number of participants = `r n_subj`

# Individual participants' global performance

```{r}
options(digits = 3)
perf <- alld %>%  group_by(subject_id) %>% summarize(group=group, prop_correct_responses = mean(correct, na.rm = TRUE), median_rt = median(reaction_time, na.rm=TRUE))
perf

p <- ggplot(perf, aes(x=prop_correct_responses, y=median_rt, colour = group, label = subject_id), show.legend = FALSE) + geom_point() + geom_text(vjust=-1) + xlim(0.88, 1.00) + ylim(450, 750)
#p <- ggplot(perf, aes(x=prop_correct_responses, y=median_rt, colour = subject_id, label = subject_id), show.legend = FALSE) + geom_point() + geom_text(vjust=-1)

p + theme(legend.position = "right")
```


# Global distribution of reaction-times on correct responses

```{r, echo=FALSE, ouput="hide", message=FALSE, warning=FALSE}
allc <- alld %>% filter(correct)
```
```{r}
ggplot(allc, aes(x = reaction_time)) + 
  geom_histogram(colour = 1, fill = "white")

```
# Data cleaning

```{r}
allc <- alld %>% filter(correct & (reaction_time < 1500))
paste("Excluding", nrow(alld) - nrow(allc), "data points out of", nrow(alld), "(", (100 * round((nrow(alld)-nrow(allc))/nrow(alld),3)), "%)", "(errors or slow responses)")
```


# Flanker congruency effet (inhibition function)

The participants must classify the direction of the central arrow.
The neighboring arrows can point toward the same direction ("congruent" condition), or the opposite direction ("incongruent" condition).
The difference in performance between the incongruent and the congruent conditions measures how well the participant can "focus" his/her attention on the central arrow, inhibiting the influence of the irrelevant flanking arrows.

## Control group (20 Ss)

```{r, warning=FALSE}
controls <- filter(allc, group=="control")

x <- ezPlot(controls, dv=reaction_time, wid=subject_id, x=run, between=.(run, flanker_congruency), between_full = .list(run, flanker_congruency), split=.list(flanker_congruency))
plot(x)
x$data
```

## Experimental group

```{r, warning=FALSE}
exp <- filter(allc, group=="experimental")
flankereffect <- exp %>% group_by(subject_id, run, flanker_congruency) %>%  summarize(mean_rt = mean(reaction_time, na.rm=TRUE))
flankereffect$grouping = interaction(flankereffect$subject_id, flankereffect$flanker_congruency)
ggplot(flankereffect, aes(x=run, y=mean_rt, colour = subject_id, label = subject_id), show.legend = FALSE) +  geom_point() + geom_line(aes(group=grouping, linetype=flanker_congruency)) + geom_text(vjust=-1)  + ylim(350, 1050)

```

```{r, warning=FALSE}
ezANOVA(allc, dv=reaction_time, wid=subject_id, between=.(flanker_congruency, run))
```

# Alerting effect


In some trials, the frame(s) of boxes flash just before the presentation of the target, alerting the participants that the target comes very soon. 
If they exploit this cue, they should be faster when there is a cue than were there is none.

```{r, warning=FALSE}
alla <-subset(allc, alerting %in% c('no_cue', 'spatial_cue'))
x <- ezPlot(alla, dv=reaction_time, wid=subject_id, x=run, between=.(run, alerting), between_full = .list(run, alerting, flanker_congruency), do_lines=TRUE, split=alerting)
plot(x)
x$data
```

```{r, warning=FALSE}
ezANOVA(alla, dv=reaction_time, wid=subject_id, between=.(alerting, run))
```

# Orienting (Cue Validity effect)

When there is a spatial cue, it can be _valid_ (cueing the actual location of target) or _invalid_ (cueing the wrong location).
If participants exploit the cue, they should be faster in the _valid_ trials than in the _invalid_ ones.



```{r, warning=FALSE}
allo = subset(allc, allc$alerting=="spatial_cue")

x <- ezPlot(allo, dv=reaction_time, wid=subject_id, x=run, between=.(run, cue_validity), between_full = .list(run, cue_validity), do_lines=TRUE, split=cue_validity)
plot(x)
x$data
```

```{r}
ezANOVA(allo, dv=reaction_time, wid=subject_id, between=.(cue_validity, run))

```

# Percentage of errors per run

```{r}
error_perrun = 100 * with(alld, 1 - tapply(correct, list(run=run, sub=subject_id), mean, na.rm=T))
t(round(error_perrun, 1))
barplot(error_perrun, beside=T)
```

```{r, echo=FALSE, ouput="hide", message=FALSE, warning=FALSE}
# Comprehensive descriptive stats

cell_stats = ddply(
	.data = alld
	, .variables = .( subject_id, run , alerting, cue_validity , flanker_congruency )
	, .fun <- function(x){
		#Compute error rate as percent.
		error_rate = (1-mean(x$correct, na.rm=TRUE))*100
		#Compute mean RT (only accurate trials).
		mean_rt = mean(x$reaction_time[x$correct==1], na.rm=TRUE)
		#Compute SD RT (only accurate trials).
		sd_rt = sd(x$reaction_time[x$correct==1], na.rm=TRUE )
		return(c(error_rate=error_rate,mean_rt=mean_rt,sd_rt=sd_rt))
	}
)

#cell_stats
```



```{r, echo=FALSE, ouput="hide", message=FALSE, warning=FALSE}
# Comprehensive descriptive stats

cell_stats = ddply(
	.data = alld
	, .variables = .( subject_id, run , alerting, cue_validity , flanker_congruency )
	, .fun <- function(x){
		#Compute error rate as percent.
		error_rate = (1-mean(x$correct, na.rm=TRUE))*100
		#Compute mean RT (only accurate trials).
		mean_rt = mean(x$reaction_time[x$correct==1], na.rm=TRUE)
		#Compute SD RT (only accurate trials).
		sd_rt = sd(x$reaction_time[x$correct==1], na.rm=TRUE )
		return(c(error_rate=error_rate,mean_rt=mean_rt,sd_rt=sd_rt))
	}
)

cell_stats
```
```{r}
flanker <- cell_stats %>% group_by(flanker_congruency, run, subject_id) %>% summarise(rt=mean(mean_rt))
ggplot(flanker, aes(run, rt)) + geom_point(aes(colour=factor(flanker_congruency)))
```

```{r}
alert <- cell_stats %>% group_by(alerting, run) %>% summarise(rt=mean(mean_rt))
ggplot(alert, aes(run, rt)) + geom_point(aes(colour=factor(alerting)))
```

